The effects of host galaxy properties on merging compact binaries detectable by LIGO

### 2016-07-12 paper/paper.tex
Pushing updated

\begin{widetext}
% POINT:  Model
To characterize how much more likely LIGO will detect coalescing binaries with different masses, we use a common and
naive estimate for the volume to which advanced LIGO is sensitive \citep[see,e.g.,][]{PSellipticals}:\footnote{For
  simplicity, in this calculation we neglect the effects of cosmology; strong field coalescence; and black hole spin;
  see \cite{popsyn-LowMetallicityImpact2c-StarTrackRevised-2014}, \cite{AstroPaper}, or \cite{RatesPaper} for more details.}
\begin{eqnarray}
V = \frac{4\pi}{3} [445 \unit{Mpc}]^3 \int p(m_1,m_2|Z) dm_1 dm_2 [ (\mc/1.2 M_\odot)^{5/6}]^3
\end{eqnarray}
where $\mc\equiv (m_1 m_2)^{3/5}/(m_1+m_2)^{1/5}$.   
This expression depends explicitly on an assumed and metallicity-dependent mass distribution for compact object
mergers, through the characteristic chirp mass $\mc_{*}(Z)\equiv [\int p(m_1,m_2|Z) dm_1 d,m_2 (\mc/1.2
  M_\odot)^{15/6}]^{5/16}$ .  We calibrate our metallicity-dependent mass distributions to metallicity-dependent binary
evolution calculations presented in \cite{popsyn-LowMetallicityImpact2-StarTrackRevised-2012} and
\cite{popsyn-LowMetallicityImpact2c-StarTrackRevised-2014}.  For neutron
stars, we adopt a fiducial neutron star mass of $1.4 M_\odot$ at all metallicities.  For BH-NS binaries, we adopt a
highly simplified model: the black hole masses are uniformly  drawn from $5 M_\odot$ to the maximum 
black hole mass $M_{max}(Z)$ allowed by
\emph{isolated} stellar evolution, as reported in prior work \citep[see,e.g.][and references
  therein]{gwastro-EventPopsynPaper-2016}.  In agreement with much more detailed prior work \cite{popsyn-LowMetallicityImpact2c-StarTrackRevised-2014}, this simplified model yields typical chirp masses for BH-NS binaries that
vary slightly as $Z/Z_\odot$ decreases, from a lower limit of $3 M_\odot$ near solar metallicity to an upper limit of
$4.3 M_\odot$ in low-metallicity environments.  Not least because the average chirp mass for BH-NS simply cannot vary
dramatically, given the functional form of $\mc$, our results do not depend sensitively on the choice of black hole mass
distribution.  Finally, for binary black holes, we assume  comparable-mass binaries
form ($m_1=m_2$), with the distribution of component masses chosen to be $\propto m_1^{-p}$ 
 between $3 M_\odot$ and $M_{max}(Z)$ and zero otherwise, adopting a fiducial exponent $p=2$; see, e.g.,
\cite{popsyn-LowMetallicityImpact2b-StarTrackRevised-2013}.    For binary black holes, this means  the
detection-weighted mass distribution [$\propto p(m_1,m_2)
\mc^{15/6}$] depends weakly on mass [$ p(m_1,m_2)
\mc^{15/6} \propto m_1^{-0.5} \delta(m_1-m_2)$.  In our simple model, the total  and chirp mass
distributions   are qualitatively consistent with detailed  binary evolution calculations \citep{popsyn-LowMetallicityImpact2b-StarTrackRevised-2013}.

@@

5-7.\footnote{This choice of exponent provides an extremely conservative assessment of the impact of low Z
\cite[see,e.g.][]{2012CQGra..29n5011O,gwastro-EventPopsynPaper-2016}. } 

@@

The close agreement between the two galaxies' binary black hole populations  is entirely a
a numerical coincdence between the specific star formation rate and metallicities in these two galaxies.  The effects described here
  are well within the scatter around the mass-metallicity relation.  Marginally different realizations of these
  histories can easily produce factors of order unity difference in galaxies with otherwise indistinguishable
  present-day properties.\editremark{do we have another example}

@@

Even when electromagnetic counterparts are available, the long delay times between binary formation and merger limit
the prospects of detailed electromagnetic followup of individual event. 
The metallicity of stars and gas adjacent to a specific merger event provides
few direct, unambiguous clues to a compact binary merger's progenitors.  
On the one hand, compact binaries are kicked by supernovae, moving substantially away from their birth position
%  2014arXiv1407.3796S : implications for nucleosynthesis
\cite{2013ApJ...776...18F,2014ARAA..52...43B}. 
On the other hand, particularly during the epoch of galaxy formation, galaxies are well-mixed: stars and adjacent gas generally do not have
similar chemical composition.  
% POINT: Host tracer gas --we are optimistically assuming the metallicity 
%  can be measured.  A seperate question is whether you can measure Z in the local environment at all.
These mixing effects have been previously recognized as an obstacle to interpreting transient event spectra.  For
example,   \cite{2010MNRAS.402.1523P} previously demonstrated that absorbing gas neighboring transient events (there, long GRBs)
would generally have high metallicity, even for low-metallicity progenitors.   
%
\citet{2010MNRAS.402.1523P} have previously used hydrodynamical simulations to demonstrate that observed ambient
metallicities (there, using damped Lyman $\alpha$ absorbers in the host) do not tightly constrain the metallicity
distribution of the progenitor; see, e.g., their Fig 3.
%

% POINT: 
Despite the challenges involved in obtaining sufficient statistics, possibly requiring third-generation instruments to
obtain sufficiently many associations, past experience suggests  host galaxy associations
will provide unique clues into the formation mechanism of compact binaries.
% 
Each host galaxy associates a merger to a unique star formation history and metallicity distribution.
With many events, these associations can potentially determine the ``response function'' for compact binaries: how often star forming gas of a given metallicity evolves into
merging compact binaries.    
%
For example, in several cases the metallicity of gas neighboring a long GRB (\citet{ 2008AJ....135.1136M}; \citet{2010AJ....140.1557L}
\editremark{links}), short GRB (\editremark{XXX}), or SN has been directly measured.
%
The precise host offset can be compared to the distribution of light and star formation  \citep{2010ApJ...708....9F}.
%
Finally,  on a host-by-host basis, the delay time between birth and merger has been constrained for long GRBs (\editremark{XXX}), short GRBs \citep{2010ApJ...725.1202L}, and
  SN Ia \citep{2011MNRAS.412.1508M}.  
%
%The future of gravitational wave host galaxy associations is bright.

>>>>

\begin{widetext}
% POINT:  Model
To characterize how much more likely LIGO will detect coalescing binaries with different masses, we use a common and
naive estimate for the volume to which advanced LIGO is sensitive \citep[see,e.g.,][]{PSellipticals}:\footnote{For
  simplicity, in this calculation we neglect the effects of cosmology; strong field coalescence; and black hole spin;
  see \cite{popsyn-LowMetallicityImpact2c-StarTrackRevised-2014}, \cite{AstroPaper}, or \cite{RatesPaper} for more details.}
\begin{eqnarray}
V = \frac{4\pi}{3} [445 \unit{Mpc}]^3 \int p(m_1,m_2|Z) dm_1 dm_2 [ (\mc/1.2 M_\odot)^{5/6}]^3
\end{eqnarray}
where $\mc\equiv (m_1 m_2)^{3/5}/(m_1+m_2)^{1/5}$.   
This expression depends explicitly on an assumed and metallicity-dependent mass distribution for compact object
mergers, through the characteristic chirp mass $\mc_{*}(Z)\equiv [\int p(m_1,m_2|Z) dm_1 d,m_2 (\mc/1.2
  M_\odot)^{15/6}]^{5/16}$ .  We calibrate our metallicity-dependent mass distributions to metallicity-dependent binary
evolution calculations presented in \cite{popsyn-LowMetallicityImpact2-StarTrackRevised-2012} and
\cite{popsyn-LowMetallicityImpact2c-StarTrackRevised-2014}.  For neutron
stars, we adopt a fiducial neutron star mass of $1.4 M_\odot$ at all metallicities.  For BH-NS binaries, we adopt a
highly simplified model: the black hole masses are uniformly  drawn from $5 M_\odot$ to the maximum 
black hole mass $M_{max}(Z)$ allowed by
\emph{isolated} stellar evolution, as reported in prior work \citep[see,e.g.][and references
  therein]{gwastro-EventPopsynPaper-2016}.  In agreement with much more detailed prior work \cite{popsyn-LowMetallicityImpact2c-StarTrackRevised-2014}, this simplified model yields typical chirp masses for BH-NS binaries that
vary slightly as $Z/Z_\odot$ decreases, from a lower limit of $3 M_\odot$ near solar metallicity to an upper limit of
$4.3 M_\odot$ in low-metallicity environments.  Not least because the average chirp mass for BH-NS simply cannot vary
dramatically, given the functional form of $\mc$, our results do not depend sensitively on the choice of black hole mass
distribution.  Finally, for binary black holes, we assume  comparable-mass binaries
form ($m_1=m_2$), with the distribution of component masses chosen to be $\propto m_1^{-p}$ 
 between $5 M_\odot$ and $M_{max}(Z)$ and zero otherwise, adopting a fiducial exponent $p=2$; see, e.g.,
\cite{popsyn-LowMetallicityImpact2b-StarTrackRevised-2013}.    For binary black holes, this means  the
detection-weighted mass distribution [$\propto p(m_1,m_2)
\mc^{15/6}$] depends weakly on mass [$ p(m_1,m_2)
\mc^{15/6} \propto m_1^{-0.5} \delta(m_1-m_2)$.  In our simple model, the total  and chirp mass
distributions   are qualitatively consistent with detailed  binary evolution calculations \citep{popsyn-LowMetallicityImpact2b-StarTrackRevised-2013}.

@@

5-7. This choice of exponent provides an extremely conservative assessment of the impact of low Z
\cite[see,e.g.,][]{2012CQGra..29n5011O,gwastro-EventPopsynPaper-2016}. 
%% Detailed calculations suggest the exponent varies
%% smoothly with metallicity, such that $d\ln \lambda/d\ln Z$ is convex.
%%  } 

@@

The close agreement between the two galaxies' binary black hole populations  is entirely a
a numerical coincdence between the specific star formation rate and metallicities in these two galaxies.  The effects described here
  are well within the scatter around the mass-metallicity relation.  Marginally different realizations of these
  histories can easily produce factors of order unity difference in galaxies with otherwise indistinguishable
  present-day properties.

@@

Even when electromagnetic counterparts are available, the long delay times between binary formation and merger limit
the prospects of detailed electromagnetic followup of individual event. 
The metallicity of stars and gas adjacent to a specific merger event provides
few direct, unambiguous clues to a compact binary merger's progenitors.  
On the one hand, compact binaries are kicked by supernovae, moving substantially away from their birth position
%  2014arXiv1407.3796S : implications for nucleosynthesis
\citep{2013ApJ...776...18F,2014ARAA..52...43B}. 
On the other hand, particularly during the epoch of galaxy formation, galaxies are well-mixed: stars and adjacent gas generally do not have
similar chemical composition.  
% POINT: Host tracer gas --we are optimistically assuming the metallicity 
%  can be measured.  A seperate question is whether you can measure Z in the local environment at all.
These mixing effects have been previously recognized as an obstacle to interpreting transient event spectra.  For
example,   \cite{2010MNRAS.402.1523P} previously demonstrated that absorbing gas neighboring transient events (there, long GRBs)
would generally have high metallicity, even for low-metallicity progenitors.   
%
\citet{2010MNRAS.402.1523P} have previously used hydrodynamical simulations to demonstrate that observed ambient
metallicities (there, using damped Lyman $\alpha$ absorbers in the host) do not tightly constrain the metallicity
distribution of the progenitor; see, e.g., their Fig 3.
%

% POINT: 
Despite the challenges involved in obtaining sufficient statistics, possibly requiring third-generation instruments to
obtain sufficiently many associations, past experience suggests  host galaxy associations
will provide unique clues into the formation mechanism of compact binaries.
% 
Each host galaxy associates a merger to a unique star formation history and metallicity distribution.
With many events, these associations can potentially determine the ``response function'' for compact binaries: how often star forming gas of a given metallicity evolves into
merging compact binaries.    
%
For example, in several cases the metallicity of gas neighboring a long GRB (\citet{ 2008AJ....135.1136M};
\citet{2010AJ....140.1557L}) has been directly measured.  % , short GRB (\editremark{XXX}), or SN has been directly measured.
%
The precise host offset can be compared to the distribution of light and star formation  \citep{2010ApJ...708....9F}.
%
Finally,  on a host-by-host basis, the delay time between birth and merger has been constrained for
% long GRBs (\editremark{XXX}), 
short GRBs \citep{2010ApJ...725.1202L}
 and
  SN Ia \citep{2011MNRAS.412.1508M}; see, e.g., the review in  \cite{2014ARAA..52...43B}.
%
%The future of gravitational wave host galaxy associations is bright.
